---
title: "ABC7 Neighborhood Safety Tracker"
# author: "John Kelly"
date: "Updated `r Sys.Date()` by the ABC7 Data Team"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(tidyverse)
library(tidycensus)
library(readxl)
library(leaflet)
library(leaflet.extras)
library(leaflet.providers)
library(sp)
library(sf)
library(htmlwidgets)
library(htmltools)

```
<link href="https://fonts.googleapis.com/css2?family=Roboto:wght@100;300;400;500;700;900&display=swap" rel="stylesheet">

```{css, echo=FALSE}

          
.title {
  font-family: roboto;
  color: #00318b;
  font-weight: 900;
  text-align: center;
  font-size: 45px;
    padding: 5px;
}

.date {
  font-family: roboto;
  color: black;
  font-weight: 300;
  text-align: center;
  font-size: 12px;
    padding: 5px;
}

select {
  margin: 0px;
  width: 175px;
  color: #00318b;
  padding: 5px 35px 5px 5px;
    font-family: roboto;
  font-size: 18px;
  font-weight: 900;
  border: 0px;
  height: 34px;
  -webkit-appearance: none;
  -moz-appearance: none;
  appearance: none;
  background: url(https://upload.wikimedia.org/wikipedia/commons/9/9d/Arrow-down.svg) 96% / 15% no-repeat #f2f2f2;
}

h1 {
  font-family: roboto;
  color: black;
  font-weight: bolder;
  text-align: left;
  font-size: 36px;
  margin-top: 0;
  margin-bottom: 0;
}

h2 {
  font-family: roboto;
  font-weight: 500;
  color: black;
  text-align: center;
  font-size: 18px;
  margin-top: 0;
  margin-bottom: 2;
}

bignumber {
  font-family: roboto;
  color: #00318b;
  font-weight: 900;
  text-align: center;
  font-size: 80px;
    line-height: 85px;
  height: 85px;
    margin-top: 0;
  margin-bottom: 0;
}

popupchatter {
  font-family: roboto;
  color: black;
  font-weight: 300;
  text-align: left;
  font-size: 12px;
  margin-top: 0;
  margin-bottom: 2;
}

chartbignumber {
  font-family: roboto;
  color: #00318b;
  font-weight: bolder;
  text-align: right;
  font-size: 36px;
    margin-top: 0;
  margin-bottom: 0;
}

subhead {
  font-family: roboto;
  color: black;
  font-weight: 700;
  text-align: left;
  font-size: 20px;
    padding: 0px;
}

body {
  color: black;
  font-family: roboto;
  font-weight: 400;
  font-size: 16px;
}

popuptitle {
  font-family: roboto;
  color: #00318b;
  font-weight: bolder;
  text-align: left;
  font-size: 20px;
    margin-top: 0;
  margin-bottom: 0;
}

h8 {
  color: #00318b;
  font-family: roboto;
  font-weight: 900;
  font-size: 16px;
}

table {
  font-family: roboto;
  border-collapse: collapse;
  width: 100%;
}

tr {
  border-bottom: thin solid #99a0a5;
}
  
td {
  text-align: right;
  padding: 6px;
}

th {
  text-align: right;
  padding: 6px;
}

   * {
      box-sizing: border-box;
   }
   .card {
      color: white;
      float: left;
      width: calc(25% - 20px);
      padding: 10px;
      border-radius: 10px;
      margin: 10px;
      height: 100%;
   }
   .card p {
   font-family: roboto;
   text-align: center;
   font-size: 12px;
  margin-bottom: 0;
   }
   .cardContainer:after {
      content: "";
      display: table;
      clear: both;
   }
   @media screen and (max-width: 650px) {
      .card {
         width: 100%;
      }

   }
```

```{r beats, include=FALSE}
# source("process_houston_police_beats.R")
```


```{r crimedata, include=FALSE}
# source("process_houston_crime_data.R")
```

<h2>ABC7's data team is tracking crime and safety across New York City and in your neighborhood.<br>You can choose which crime to examine: <select onchange="window.location=this.value">
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker.html">Homicides</option>
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker_SexualAssaults.html">Sexual Assaults</option>
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker_AutoThefts.html">Auto Thefts</option>
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker_Assaults.html">Assaults</option>
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker_Thefts.html">Thefts</option>
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker_Burglaries.html">Burglaries</option>
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker_Robberies.html">Robberies</option>
<option value="https://abcotvdata.github.io/safetytracker_newyork/New_York_City_Safety_Tracker_DrugCrimes.html">Drug Crimes</option>
</select></h2>

<div class="cardContainer">
<div class="card" style="background-color:#99a0a5;">
<p>So far this year<br>
<bignumber>`r prettyNum(murders_city$total22, big.mark=",")`<br></bignumber><p>
people have been murdered.</p>
</div>
<div class="card" style="background-color:#99a0a5;">
<p>Over the last 12 momths,<br>
<bignumber>`r prettyNum(murders_city$last12mos, big.mark=",")`<br></bignumber>
<p>people have been killed.</p>
</div>
<div class="card" style="background-color:#99a0a5;">
<p>The yearly average is<br>
<bignumber>`r prettyNum(round(murders_city$avg_prior3years,0), big.mark=",")`<br></bignumber><p>
from 2019 to 2021.</p>
</div>
<div class="card" style="background-color:#99a0a5;">
<p>The murder rate is <br>
<bignumber>`r murders_city$rate_last12`<br></bignumber><p>
per 100K people the last year.</p>
</div>
</div>
<br>
About three years ago, the murder rate in New York City was <h8>`r murders_city$rate19`</h8> per 100,000 residents - or just slightly less than the rate of deaths in motor vehicle crashes in Texas (12.5).

During the pandemic years, the risk of homicide in Houston rose above the risk of being killed in a traffic crash. The average murder rate for the three-year period from 2019 to 2022 was <h8>`r murders_city$rate_prior3years`</h8> and last year it was <h8>`r murders_city$rate21`</h8>.

The situation, of course, is not the same neighborhood to neighborhood.

About <h8>`r round(pop_murdersdown/2304580*100,0)`%</h8> of city residents live in neighborhoods where the murder rate over the last 12 months is down or about the same as the three-year average. 

In fact, <h8>`r round(pop_murderszero/2304580*100,0)`%</h8> of people live in neighborhoods where there were no homicides at all over the last 12 months.

But, about <h8>`r round(pop_murdersup/2304580*100,0)`%</h8> of Houston residents live in neighborhoods where the murder rate is up more than 20% in the last 12 months when compared to the three-year average from 2019 to 2021.

And, <h8>`r round(pop_murdersmonthly/2304580*100,0)`%</h8> of us now live in neighborhoods where there was at least one murder every month over the last year.

Across the city the last 12 months, there were about <h8>`r round(murders_city$last12mos/52,1)`</h8> homicides every week. In 2019, that number was <h8>`r round(murders_city$total19/52,1)`</h8> per week

<subhead>Houston homicides neighborhood by neighborhood</subhead>

ABC7 looked at the New York Police Department's data, beat by beat across the city and is providing this look at the murder trends by neighborhood for each of the last four years in the map below. The color-coding of the map is based on the murder rate over 2019, 2020 and 2021 combined, according to ABC13's analysis of data the city police department publishes monthly.

```{r cars, echo=FALSE,out.width='100%',out.height='600'}
# MURDER MAP

# Set bins for numbers of crimes for murders map
murderpal <- colorQuantile(c("#667f99",
                             "#00318b",
                             "#0058f6",
                             "#ffba00"), murders_beat$rate_prior3years, n = 4, na.color = NA)

# Create quick labels for murders map
murderlabel <- paste(sep="",
"<table>
<popuptitle>Houston P.D. Beat #",
murders_beat$beat,
"</popuptitle><br><popupchatter>Area neighborhoods and landmarks include ",
murders_beat$placename,
". Estimated Population: ",
prettyNum(murders_beat$population, big.mark=','),
"</popupchatter><br>
      <tr>
				<th></th>
				<th>Total</th>
				<th>Rate</th>
				<th>Trend</th>
			</tr>
			<tr>
				<td>2019</td>
				<td>",
murders_beat$total19,
"</td>
				<td>",
murders_beat$rate19,
"</td>
				<!-- Insert cell to cross/span 
				multi rows -->
				<td rowspan='5'>",
ifelse(murders_beat$inc_prior3yearavgtolast12>=0,"Last 12<br>months up","Last 12<br>months down"),
"<br><chartbignumber>",
ifelse(murders_beat$inc_prior3yearavgtolast12>=0,murders_beat$inc_prior3yearavgtolast12,substring(murders_beat$inc_prior3yearavgtolast12,2)),
				"%</chartbignumber><br>vs. 3-year<br>average</td>
			</tr>
			<tr>
				<td>2020</td>
				<td>",
murders_beat$total20,
"</td>
				<td>",
murders_beat$rate20,
"</td>
			</tr>
						<tr>
				<td>2021</td>
				<td>",
murders_beat$total21,
"</td>
				<td>",
murders_beat$rate21,
"</td>
			</tr>
						<tr>
				<td>Last 12 months</td>
				<td>",
murders_beat$last12mos,
"</td>
				<td>",
murders_beat$rate_last12,
"</td>
			</tr>
						<tr>
				<td>3-year Average</td>
				<td>",
murders_beat$avg_prior3years,
"</td>
				<td>",
murders_beat$rate_prior3years,
"</td>
			</tr>
</table>")

# Creating police beats map for types of crimes
houston_murder_map <- leaflet(murders_beat, options = leafletOptions(zoomControl = FALSE)) %>%
  htmlwidgets::onRender("function(el, x) {
L.control.zoom({ position: 'topright' }).addTo(this)
}") %>%
  setView(-95.4, 29.75, zoom = 12) %>% 
  addProviderTiles(provider = "Stamen.TonerLite") %>%
  addPolygons(color = "white", 
              popup = murderlabel, 
              weight = 1, 
              smoothFactor = 0.5,
              opacity = 0.6, 
              fillOpacity = 0.5,
              fillColor = ~murderpal(rate_prior3years)) %>% 
  addLegend(opacity = 0.6,
            values = murders_beat$rate_prior3years, 
            colors = c("#667f99",
                       "#00318b",
                       "#0058f6",
                       "#ffba00"),
            labels=c("Lowest Rate", 
                     "Lower Rate", 
                     "Higher Rate", 
                     "Highest Rate"),
            position = "bottomleft", 
            title = "Homicides<br>Per 100K people<br>") 

houston_murder_map
```

However, in the vast majority of neighborhoods in the city, the rate is much lower. In 17 of 118 of the Houston Police Department's, the murder rate was zero last year. Filler, filler, filler.
<br>
<br>
<!-- <subhead>Homicides month by month, 2019 through last month</subhead> -->

<iframe title="Homicides month by month, 2019 through last month" aria-label="Column Chart" id="datawrapper-chart-3NHzG" src="https://datawrapper.dwcdn.net/3NHzG/5/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="400" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>
  <br>
<br>
<!-- <subhead>Risk of being murdered in Houston vs. other causes of death in Texas</subhead> -->

<iframe title="" aria-label="Dot Plot" id="datawrapper-chart-9SHHa" src="https://datawrapper.dwcdn.net/9SHHa/5/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="146" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>
<br>
<br>
<!-- <subhead>What kinds of places do murders happen in Houston</subhead> -->

<iframe title="What kinds of places do murders happen in Houston"  aria-label="Table" id="datawrapper-chart-xI2y0" src="https://datawrapper.dwcdn.net/xI2y0/10/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="1239" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>
<br>
<br>
<!-- <subhead>What time of day do murders happen?</subhead> -->

<iframe title="What hour of day do murders happen" aria-label="Column Chart" id="datawrapper-chart-coxAA" src="https://datawrapper.dwcdn.net/coxAA/7/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="400" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>
